Machine learning enhanced empirical potentials for metals and alloys

### 2020-03-17 doc/papers/ml-eam/manuscript.tex
This is the final version for npj.

\author{Hai-Feng Song}
\email{song_haifeng@iapcm.ac.cn}
\affiliation{
    Institute of Applied Physics and Computational Math, 
    Beijing 100088, China}
\affiliation{
    CAEP Software Center for High Performance Numerical Simulation, 
    Beijing 100088, China}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Abstract
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{abstract}
Interatomic potential or force-field plays a vital role in simulating materials. 
Empirical potentials like the embedded atom method (EAM) and its variant 
angular-dependent potential (ADP) have proven successful in modeling many 
metals qualitatively. In the past few years, machine learning has became the 
key approach for modeling interatomic interactions. Powered by big data and 
efficient optimizers, machine learning interatomic potentials can generally 
approximate to the first-principles accuracy. In this works, we successfully 
integrated machine learning interatomic potential with EAM and ADP. Machine 
learning can significantly boost their performances. For energy and force 
predictions, machine leaned EAM and ADP can be almost as accurate as the 
spectral neighbor analysis potential (SNAP) on the fcc Ni, bcc Mo and Mo-Ni 
alloy systems, while the molecular dynamics simulations performed by the later 
approach is nealy 1000 times slower. Machine learned EAM and ADP can also give 
first-principle accuracy key material properties, such as elastic constants, 
melting temperatures and surface energies. Our results suggest a new and 
systematic route for developing machine learning interatomic potentials. 
\end{abstract}

@@

@@

We firstly use the two elementary datasets (Ni and Mo) to show our approach. These 
two datasets has 461 and 284 structures, respectively. 61 and 34 were randomly 
selected to form the test subsets. In these two experiments, 
$\chi_{\mathrm{f}}$, $\chi_{\mathrm{rose}}$ and $\chi_{\mathrm{elastic}}$ were 
set to 1, 3 and 0.05 respectively. The batch size was fixed to 25.

@@

Figure \ref{fig:Ni_eam} compares the $\rho(r)$, $F(\rho)$, $\phi(r)$, 
$\mu(r)$ and $\omega(r)$ functions. The differences between ML-EAM and ML-ADP 
$\rho(r)$, $\phi(r)$ and $F(\rho)$ are quite small. For ML-ADP, both the dipole 
and the quadrupole terms converge to zero at $\sim 4$ \AA. The dipole term gives
the major contribution.

@@

For the Mo dataset, the situation is quite different. The original Zjw04 has 
extremely large MAEs: the energy \textbf{rMAE} is $\sim 98$ meV/atom and the 
force MAE is more than 0.40 eV/\AA. Our method can significantly reduce the 
error: ML-EAM can achieve test MAEs of 23.8 meV/atom and 0.26 eV/\AA and overall
MAEs of 26.7 meV/atom and 0.30 eV/\AA. We also optimized an ADP potential for 
this dataset. Interestingly, ML-ADP gets better results: 20.6 meV/atom and 
0.24 eV\AA on the test set and 18.7 meV/atom and 0.29 eV/\AA on the entire 
dataset. On the contrast, the SNAP method (13.2 meV/atom, 0.25 eV/\AA) behaves 
slightly better than ML-ADP. But considering the computational cost, the 
performace of ML-ADP is totally acceptable.

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Mo - EAM/ADP - Plot
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}[htp]
\centering
\includegraphics[scale=0.6]{figures/Mo_eam_adp.png}
\caption{\label{fig:Mo_eam_adp} $\rho(r)$, $F(\rho)$, $\phi(r)$ of the original 
Zjw04 EAM, ML-EAM and ML-ADP. \textbf{(d)} shows the $\mu(r)$ and $\omega(r)$ of 
ML-ADP.}
\end{figure}

>>>>

\author{Hai-Feng Song}
\email{song\_haifeng@iapcm.ac.cn}
\affiliation{
    Institute of Applied Physics and Computational Math, 
    Beijing 100088, China}
\affiliation{
    CAEP Software Center for High Performance Numerical Simulation, 
    Beijing 100088, China}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Abstract
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{abstract}
Interatomic potential or force-field plays a vital role in simulating materials. 
Empirical potentials like the embedded atom method (EAM) and its variant 
angular-dependent potential (ADP) have proven successful in modeling many 
metals qualitatively. In the past few years, machine learning has became the 
key approach for modeling interatomic interactions. Powered by big data and 
efficient optimizers, machine learning interatomic potentials can generally 
approximate to the first-principles accuracy. In this works, we successfully 
developed a route to express EAM and ADP within machine learning framework in 
highly-vectorizable form and further enhanced the potentials with physical 
constraints. Machine learning can significantly boost their performances with 
few training data. 
For energy and force 
predictions, machine leaned EAM and ADP can be almost as accurate as the 
spectral neighbor analysis potential (SNAP) on the fcc Ni, bcc Mo and Mo-Ni 
alloy systems, while the molecular dynamics simulations performed by the later 
approach is nealy 1000 times slower. Machine learned EAM and ADP can also give 
first-principle accuracy key material properties, such as elastic constants, 
melting temperatures and surface energies. Our results suggest a new and 
systematic route for developing machine learning interatomic potentials. 
\end{abstract}

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Mo - EAM/ADP - Plot
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}[htp]
\centering
\includegraphics[scale=0.6]{figures/Mo_eam_adp.png}
\caption{\label{fig:Mo_eam_adp} $\rho(r)$, $F(\rho)$, $\phi(r)$ of the original 
Zjw04 EAM, ML-EAM and ML-ADP for elementary Mo. \textbf{(d)} shows the $\mu(r)$ 
and $\omega(r)$ of ML-ADP.}
\end{figure}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Table: Ni / Mo MAEs
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{table}
\centering
\begin{tabular}{lccccc}
\hline
            &    & EAM\cite{ZJW2} & SNAP\cite{SNAP} & ML-EAM & ML-ADP \\
\hline
Energy     & Ni & 10.6           & 1.2             & 3.9    &        \\
(meV/atom) & Mo & 58.9           & 13.2            & 23.8   & 18.7   \\
\hline
Force      & Ni & 0.06           & 0.05            & 0.05   &        \\
(eV/\AA)   & Mo & 0.31           & 0.25            & 0.30   & 0.29   \\
\hline
\end{tabular}
\caption{\label{table:Mo_Ni_mae} Comparisons of the energy and force MAEs of 
elementary Ni and Mo. 
}
\end{table}

@@

We firstly use the two elementary datasets (Ni and Mo) to show our approach. These 
two datasets has 461 and 284 structures, respectively. 61 and 34 structures were 
randomly selected to form the test subsets. In these two experiments, 
$\chi_{\mathrm{f}}$, $\chi_{\mathrm{rose}}$ and $\chi_{\mathrm{elastic}}$ were 
set to 1, 3 and 0.05 respectively. The batch size was fixed to 25.

@@

Figure \ref{fig:Ni_eam} compares the $\rho(r)$, $F(\rho)$, $\phi(r)$, 
$\mu(r)$ and $\omega(r)$ functions. The differences between ML-EAM and ML-ADP 
$\rho(r)$, $\phi(r)$ and $F(\rho)$ are quite small.

@@

For the Mo dataset, the situation is quite different. The original Zjw04 has 
extremely large MAEs: the energy \textbf{rMAE} is $\sim 59$ meV/atom and the 
force MAE is more than 0.31 eV/\AA. Our method can significantly reduce the 
error: ML-EAM can achieve test MAEs of 23.8 meV/atom and 0.26 eV/\AA and overall
MAEs of 26.7 meV/atom and 0.30 eV/\AA. We also optimized an ADP potential for 
this dataset. Interestingly, ML-ADP gets better results: 20.6 meV/atom and 
0.24 eV\AA on the test set and 18.7 meV/atom and 0.29 eV/\AA on the entire 
dataset. On the contrast, the SNAP method (13.2 meV/atom, 0.25 eV/\AA) behaves 
slightly better than ML-ADP. But considering the computational cost, the 
performace of ML-ADP is totally acceptable. Figure \ref{fig:Mo_eam_adp} 
demonstrates the potential functions. The $\rho(r)$, $F(\rho)$ and $\phi(r)$ 
functions of ML-EAM and ML-ADP are quite similar. The additional accuracy of 
ML-ADP mainly comes from the dipole function $\mu(r)$.

@@
